---
title: "Air Pollution Analysis with Satellite Data"
author: "satellite package developers"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Air Pollution Analysis with Satellite Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `satellite` package now includes comprehensive air pollution monitoring capabilities through integration with major satellite atmospheric datasets. This vignette demonstrates how to download, process, and analyze air pollution data from Sentinel-5P TROPOMI and MODIS atmospheric products.

## Supported Platforms and Pollutants

### Sentinel-5P TROPOMI
- **Nitrogen Dioxide (NO2)**: Tropospheric and stratospheric columns
- **Sulfur Dioxide (SO2)**: Total vertical columns  
- **Carbon Monoxide (CO)**: Total column measurements
- **Ozone (O3)**: Total vertical columns
- **Formaldehyde (HCHO)**: Tropospheric columns
- **Methane (CH4)**: Mixing ratios
- **Aerosol Index**: UV aerosol index and layer height

### MODIS Atmospheric Products  
- **Aerosol Optical Depth (AOD)**: Terra/Aqua aerosol products
- **Fire Data**: Active fire detections and thermal anomalies
- **Atmospheric Parameters**: Cloud properties and water vapor

## Downloading Sentinel-5P Data

```{r eval=FALSE}
library(satellite)

# Download NO2 data for a specific region and time period
no2_result <- downloadSentinel5P(
  product = "L2__NO2___",
  start_date = "2023-06-01", 
  end_date = "2023-06-07",
  bbox = c(-74.5, 40.0, -73.5, 41.0), # NYC area
  cloud_cover = 30,
  output_dir = "./s5p_data",
  source = "copernicus",
  username = "your_username",
  password = "your_password",
  max_results = 10
)

print(no2_result$metadata)
```

## Processing Sentinel-5P Data

```{r eval=FALSE}
# Process downloaded NetCDF files
sat_no2 <- processSentinel5P(
  files = no2_result$downloaded_files,
  parameters = c("nitrogendioxide_tropospheric_column"),
  quality_filter = TRUE,
  quality_threshold = 0.7,
  cloud_filter = TRUE,
  cloud_threshold = 0.3,
  unit_conversion = TRUE
)

# View the processed satellite object
print(sat_no2)
```

## Downloading MODIS Atmospheric Data

```{r eval=FALSE}
# Download MODIS aerosol data
aod_result <- downloadMODISPollution(
  product = "MOD04_L2",
  start_date = "2023-06-01",
  end_date = "2023-06-07", 
  bbox = c(-125, 30, -110, 45), # California
  output_dir = "./modis_data",
  source = "earthdata",
  username = "your_nasa_username",
  password = "your_nasa_password",
  max_files = 20
)

# Process MODIS files
sat_aod <- processMODISPollution(
  files = aod_result$downloaded_files,
  parameters = c("Optical_Depth_Land_And_Ocean", "Fine_Mode_Fraction_Ocean"),
  quality_filter = TRUE,
  scale_factors = TRUE,
  mask_fill_values = TRUE
)
```

## Statistical Analysis

```{r eval=FALSE}
# Calculate comprehensive NO2 statistics
no2_stats <- calcPollutionStats(
  sat_no2,
  pollutant = "NO2",
  temporal_analysis = TRUE,
  spatial_analysis = TRUE,
  exceedance_analysis = TRUE
)

# View summary statistics
print(no2_stats$summary_stats)

# Check for exceedances of air quality standards
print(no2_stats$exceedances)
```

## Time Series Visualization

```{r eval=FALSE}
# Create time series plot with trend analysis
no2_plot <- plotPollutionTimeSeries(
  sat_no2,
  pollutant = "NO2", 
  trend_line = TRUE,
  anomalies = TRUE,
  aggregation = "monthly",
  title = "NYC NO2 Trends"
)

print(no2_plot)

# Plot aerosol time series
aod_plot <- plotPollutionTimeSeries(
  sat_aod,
  pollutant = "AOD",
  trend_line = TRUE,
  seasonal = TRUE,
  title = "California Aerosol Trends"
)

print(aod_plot)
```

## Data Quality and Validation

The package includes comprehensive quality assurance capabilities:

### Sentinel-5P Quality Filtering
- **QA Values**: 0-1 scale quality indicators
- **Cloud Filtering**: Remove cloudy pixels
- **Retrieval Quality**: Check algorithm convergence
- **Surface Conditions**: Account for surface albedo effects

### MODIS Quality Filtering  
- **Land/Ocean Flags**: Separate quality indicators for different surfaces
- **Algorithm Quality**: Retrieval confidence levels
- **Cloud Screening**: Multiple cloud detection methods
- **Geometric Quality**: View angle and solar geometry filters

## Air Quality Standards Integration

The package includes WHO and EPA air quality guidelines:

```{r eval=FALSE}
# Custom thresholds for exceedance analysis
custom_thresholds <- list(
  NO2 = list(
    annual_who = 10,      # µg/m³ (WHO 2021)
    annual_epa = 53,      # µg/m³ (EPA)
    hourly_who = 200      # µg/m³ (WHO)
  ),
  SO2 = list(
    daily_who = 40,       # µg/m³ (WHO)
    hourly_epa = 196      # µg/m³ (EPA)
  )
)

# Apply custom thresholds
stats_custom <- calcPollutionStats(
  sat_no2,
  pollutant = "NO2",
  exceedance_analysis = TRUE,
  thresholds = custom_thresholds$NO2
)
```

## Example: COVID-19 Impact Analysis

```{r eval=FALSE}
# Download pre-COVID data
pre_covid <- downloadSentinel5P(
  product = "L2__NO2___",
  start_date = "2019-03-01",
  end_date = "2019-05-31",
  bbox = c(-74.5, 40.0, -73.5, 41.0)
)

# Download COVID-period data  
covid_period <- downloadSentinel5P(
  product = "L2__NO2___", 
  start_date = "2020-03-01",
  end_date = "2020-05-31",
  bbox = c(-74.5, 40.0, -73.5, 41.0)
)

# Process both datasets
sat_pre <- processSentinel5P(pre_covid$downloaded_files)
sat_covid <- processSentinel5P(covid_period$downloaded_files)

# Calculate change statistics
change_stats <- calcPollutionStats(
  list(pre_covid = sat_pre, covid = sat_covid),
  pollutant = "NO2",
  temporal_analysis = TRUE
)

# Visualize the change
comparison_plot <- plotPollutionTimeSeries(
  list(pre_covid = sat_pre, covid = sat_covid),
  pollutant = "NO2",
  title = "NYC NO2: Pre-COVID vs COVID-19 Lockdown"
)
```

## References

1. Veefkind, J. P., et al. (2012). TROPOMI on the ESA Sentinel-5 Precursor. Remote sensing of environment, 120, 70-83.

2. Levy, R. C., et al. (2013). The Collection 6 MODIS aerosol products over land and ocean. Atmospheric Measurement Techniques, 6(11), 2989-3034.

3. WHO (2021). WHO global air quality guidelines: particulate matter (PM2.5 and PM10), ozone, nitrogen dioxide, sulfur dioxide and carbon monoxide. World Health Organization.

## See Also

- `?downloadSentinel5P` for detailed parameter descriptions
- `?processSentinel5P` for processing options
- `?calcPollutionStats` for statistical analysis capabilities
- `?plotPollutionTimeSeries` for visualization options
